---
title: "R Notebook"
output: html_notebook
---

Loading libraries 

```{r}
library(tidyverse)
library(here)
```

```{r}
here::here()
```

Reading in clean data

```{r}
clean_decathlon <- read_csv(here("clean_data", "cleaned_decathlon_data.csv"))

```

Viewing clean data

```{r}
view(clean_decathlon)

dim(clean_decathlon)

```

Finding which athlete had the longest long jump in the data

```{r}
longest_long_jump <- clean_decathlon %>%
  filter(event == "long_jump") %>% #filter out all events that are not long_jump
  arrange(desc(distance_or_time)) %>% # order the distance_or_time values in descending order
  head(1) # return the top row which has the highest value

longest_long_jump

```

Clay is the athlete with the longest long jump distance


Finding the average 100m time for each competition

```{r}
avg_100m_time <- clean_decathlon %>%
  filter(event == "100m") %>% #filter out all events that are not 100m
  group_by(competition) %>% #group the data by competition - decastar and olympics
  summarise(avg_100m = round(mean(distance_or_time),
                             digits = 1)) 
# create new column calculating mean (average) of all 100m distance_or_time values & round to one decimal place

avg_100m_time

```

Average 100m time for Decastar is 11.2 and for the Olympics its 10.9 - rounded to one decimal place as this is standard in athletics


Finding the highest total points in dataset (across both competitions)

```{r}
highest_points_total <- clean_decathlon %>%
  select(athlete, points) %>% # focus on athlete and points columns
  arrange(desc(points)) %>% # organise them in order of points (descending)
  head(1) #return the top row which has the highest value

highest_points_total

```

Sebrle had the highest points total across both competitions - 8893


The top 3 shot-put scores (distance) for each competition

```{r}
shot_put_scores_top_3 <- clean_decathlon %>%
  filter(event == "shot_put") %>% ##filter out all events that are not shot put
  group_by(competition) %>% # group the data by competition - decastar and olympics
  filter(between(rank, 1, 3)) # only focus on the rows of the top 3 competitors (ranked between 1 and 3)

shot_put_scores_top_3


```

shot_put_scores_top_3:

decastar - 14.83, 14.26, 14.77
olympics - 16.36, 15.23, 15.93


Returning the average points for competitors above and below 50s

```{r}
avg_400m_group_above_and_below_50s <- clean_decathlon %>%
  filter(event == "400m") %>% #filter out all events that are not shot put
  mutate(above_50s = distance_or_time > 50) %>% #create a new logical column that specifies if the score is above 50
  group_by(above_50s) %>% # group by above_50s column
  summarise(avg_points = round(mean(points),
                             digits = 2)) # calculate average points for those above 50s(true) and below 50s (false), rounded to 2 decimal places
  

avg_400m_group_above_and_below_50s


```

average points for those below 50s was 8120.48 and above 50s was 7727.17



















































